Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities in depth

### 2017-08-14 manuscript/manuscript.tex
Changes in manuscript

@@

For practical purposes we are going to obtain the gravity potential for two very different functions in the perspective of their variation in $r'$: a linear, an exponential and a discontinuous density function.

@@

Finally, we set the acceptable value of $D$ at which the error of the numerical approximation is lower than 10$^{-1}\%$.

The tests were only performed on these three magnitudes because the other components of the gravity gradient and Marussi tensor are equal to zero due to the rotational symmetries of the spherical shell.

@@

For the linear density case, we can observe that the calculation of the potential and the $g_z$ component of its gradient needs a value of $D=1$ and $D=2$, respectively, in order to ensure an accuracy of 10$^{-1}\%$.
On the other hand, the $g_{zz}$ computations show a noticeable difference between the thin and the thick shell: the first one needs a value of $D=8$ while the later only a $D=3.5$.
We will keep the more conservative result of $D=8$ in case of computing the $g_{zz}$ tensor component in case of a linear density tesseroid.

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Dlinear-thin-differences.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the linear density defined on equation \ref{eq:density-linear}. The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$. If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-linear-thin}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Dlinear-thick-differences.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the linear density defined on equation \ref{eq:density-linear}. The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$. If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-linear-thick}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Dexp-shifted-thin-differences.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the exponential density defined on equation \ref{eq:density-exp}. The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$. If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-exp-thin}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Dexp-shifted-thick-differences.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the exponential density defined on equation \ref{eq:density-exp}. The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$. If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-exp-thick}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Dexp-power-differences.pdf}
\caption{
    Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for (a) a thin spherical shell of 1km of thickness and (b) a thick sphericall shell of 35km of thickness, both with the exponential density defined on equation \ref{eq:density-exp} with different values of $b$. Each computation was performed on the big grid described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$. Each curve corresponds to a value of $b$. If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-exp-power}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Ddiscontinuous-symmetric-differences.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous} and $R_c$ as the centre radius of the shell. The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$. If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-discont-sym}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Ddiscontinuous-asymmetric-differences.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous} and $R_c=(R_1 + R_2)/3$. The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$. If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-discont-asym}
\end{figure}

>>>>

\todo{corregir esto y explicar mejor}

@@

For practical purposes we are going to obtain the gravity potential for three different functions in the perspective of their variation in $r'$: a linear, an exponential and a discontinuous density function.

@@

Finally, we set the acceptable value of $D$ at which the error of the numerical approximation is lower than 0.1\%.

The tests were only performed on these three magnitudes because the other components of the gravity gradient and Marussi tensor are equal to zero due to the rotational symmetries of the spherical shell. \todo{corregir esto, los elementos diagonales del tensor fueron testeados, pero no lo mostramos pq son muy similares al gzz}

@@

For the linear density case, we can observe that the calculation of the potential and the $g_z$ component of its gradient needs a value of $D=1$ and $D=2$, respectively, in order to ensure an accuracy of 0.1\%.
On the other hand, the $g_{zz}$ computations show a noticeable difference between the thin and the thick shell: the first one needs a value of $D=8$ while the later only a $D=3.5$.
We will keep the more conservative result of $D=8$ in case of computing the $g_{zz}$ tensor component in case of a linear density tesseroid.

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Dlinear-thin-differences.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the linear density defined on equation \ref{eq:density-linear}. The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-linear-thin}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Dlinear-thick-differences.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the linear density defined on equation \ref{eq:density-linear}. The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-linear-thick}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Dexp-shifted-thin-differences.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the exponential density defined on equation \ref{eq:density-exp}. The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-exp-thin}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Dexp-shifted-thick-differences.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the exponential density defined on equation \ref{eq:density-exp}. The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-exp-thick}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Dexp-power-differences.pdf}
\caption{
    Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for (a) a thin spherical shell of 1km of thickness and (b) a thick sphericall shell of 35km of thickness, both with the exponential density defined on equation \ref{eq:density-exp} with different values of $b$. Each computation was performed on the big grid described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$. Each curve corresponds to a value of $b$. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-exp-power}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Ddiscontinuous-symmetric-differences.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous} and $R_c$ as the centre radius of the shell. The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-discont-sym}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Ddiscontinuous-asymmetric-differences.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous} and $R_c=(R_1 + R_2)/3$. The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-discont-asym}
\end{figure}
